---
title: "R Tip of the Day"
subtitle: "The ggpattern package"
author: "Jessica Schmidt"
date: "November 20, 2025"
format:
  revealjs: 
    slide-number: true
    chalkboard: 
      buttons: true
    preview-links: auto
editor_options: 
  chunk_output_type: console
---

```{r}
library(tidyverse)
library(ggpattern)
library(gganimate)
penguins <- palmerpenguins::penguins
```


## What is ggpattern?

ggpattern is an extension of ggplot 2 that lets you add patterns to your graphs.

::: incremental
-   Patterns are a new optional aesthetic in ggplot
-   New functions: geom_bar_pattern(), geom_col_pattern(), etc
-   Both built-in and custom patterns
:::

## Penguin Patterns {auto-animate=true}

```{.r code-line-numbers="3-4"}
penguins |>
  ggplot(aes(x = species)) +
  geom_bar_pattern(aes(pattern_fill = species), 
                   pattern = 'crosshatch',
                   fill = "white",
                   color = "black")
```

## Penguin Patterns {auto-animate=true}

```{r}
penguins |>
  ggplot(aes(x = species)) +
  geom_bar_pattern(aes(pattern_fill = species), 
                   pattern = 'crosshatch',
                   fill = "white",
                   color = "black")
```

## Different Pattern Options - Stripe

```{r}
penguins |>
  ggplot(aes(x = species)) +
  geom_bar_pattern(aes(pattern_fill = species), 
                   pattern = 'stripe',
                   fill = "white",
                   color = "black")
```

## Different Pattern Options - Circle

```{r}
penguins |>
  ggplot(aes(x = species)) +
  geom_bar_pattern(aes(pattern_fill = species), 
                   pattern = 'circle',
                   fill = "white",
                   color = "black")
```


## Different Pattern Options - Gradient

```{r}
penguins |>
  ggplot(aes(x = species)) +
  geom_bar_pattern(aes(pattern_fill = species), 
                   pattern = 'gradient',
                   pattern_alpha = 0.8,
                   fill = "white",
                   color = "black")
```

## Different Pattern Options - Plasma

```{r}
penguins |>
  ggplot(aes(x = species)) +
  geom_bar_pattern(aes(pattern_fill = species), 
                   pattern = 'plasma',
                   fill = "white",
                   color = "black")
```

## Different Pattern Options - Images

```{r}
penguins |>
  ggplot(aes(x = species)) +
  geom_bar_pattern(aes(pattern_fill = species), 
                   pattern = 'placeholder',
                   pattern_type = 'bear', show.legend = FALSE)
```

## Different Pattern Options - Custom Patterns

```{.r code-line-numbers="3-7"}
penguins |>
  ggplot(aes(x = species)) +
  geom_bar_pattern(aes(pattern_fill = species), 
                   pattern = 'stripe',
                   pattern_alpha = 0.5,
                   pattern_angle = 90,
                   pattern_color = "blue",
                   fill = "white",
                   color = "black")
```

## Different Pattern Options - Custom Patterns

```{r}
penguins |>
  ggplot(aes(x = species)) +
  geom_bar_pattern(aes(pattern_fill = species), 
                   pattern = 'stripe',
                   pattern_alpha = 0.5,
                   pattern_angle = 90,
                   pattern_color = "blue",
                   fill = "white",
                   color = "black")
```

## Different Plots - geom_boxplot_pattern()

```{r}
penguins |>
  ggplot(aes(x = body_mass_g, y = species)) +
  geom_boxplot_pattern(aes(pattern = species))
```

## Different Plots - geom_violin_pattern()

```{r}
penguins |>
  ggplot(aes(x = body_mass_g, y = species)) +
  geom_violin_pattern(aes(pattern = species, pattern_fill = species, fill = species))
```

## Different Plots - geom_density_pattern()

```{r}
penguins |>
  ggplot(aes(x = body_mass_g)) +
  geom_density_pattern(fill = "pink", pattern_fill = "lightgreen")
```


## Different Plots - geom_histogram_pattern()

```{r}
penguins |>
  ggplot(aes(x = body_mass_g)) +
  geom_histogram_pattern(pattern = 'crosshatch')
```


## Why do we care about patterns? {auto-animate="true"}

Patterns can make graphs more interesting, informative, and accessible. 

::: incremental
-   Patterns may make a plot more clear or exciting to look at
-   Custom patterns can emphasize parts of the data
-   Patterns are another way to make charts colorblind friendly
:::
